---
title: "Spatial Prediction via Spatial Weibull Model (SWM)"
author: "Authors: Se Yoon Lee and Bani Mallick"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**_NOTE:_** Objective of this code is to predict oil production rate trajectory at a new testing location of a shale oil well in the Eagle Ford Shale of South Texas


Load libraries/R functions/Dataset
```{r include=FALSE}
library("dplyr")
library("tidyr")
library("ggplot2")
library("maps") ; Sys.getenv("R_LIBS_USER")

setwd("-/GitHub/SWM/R codes")
source("SWM.R")

  setwd("-/GitHub/SWM/Dataset/")
  load("Production_Results_Eagle_Ford.RData")
  
  # Set number of testing wells
  N.test = 36
  N = 360 - N.test # number of training wells   
  
  Y = log(Production_Results_Eagle_Ford$P[1:N,])
  X = scale(Production_Results_Eagle_Ford$X, center = TRUE, scale = TRUE)[1:N,]
  Loc = Production_Results_Eagle_Ford$Loc[1:N,]
  
  te.Y = log(Production_Results_Eagle_Ford$P[-c(1:N),])
  te.X = scale(Production_Results_Eagle_Ford$X, center = TRUE, scale = TRUE)[-c(1:N),]
  te.Loc = Production_Results_Eagle_Ford$Loc[-c(1:N),]
```

Locations of training and testing wells

```{r echo=FALSE}
par(mfrow = c(1,1))
map('state', 'texas',fill = TRUE,col = "grey",
    xlim = c(-100.5,-96),exact = TRUE,
    ylim = c(28,30))
title("Locations of training wells (black dots) and testing wells (red triangles).")
points( x = Loc[,2], y = Loc[,1],  col = "black", pch = 20)
points( x = te.Loc[,2], y = te.Loc[,1],  col = "red", pch = 17)
```

Production trajectories of training and testing wells
```{r echo=FALSE}
y = function(i){ 
    temp.y = as.numeric(log(Production_Results_Eagle_Ford$P[i,]))
    start.i = sum(as.numeric(is.na(temp.y))) + 1
    y = c(temp.y[start.i:ncol(Y)])
    return(y)
}
temp_dataset = cbind(Production_Results_Eagle_Ford$API10,Production_Results_Eagle_Ford$P)
df_1 = gather(temp_dataset,"t","production",2:79)
df_2 = mutate(df_1, API10_char = paste("API",API10) )
df_3 = df_2 %>% 
  filter(production > 0) %>% 
  group_by(API10_char) %>% 
  summarise(n = n(), mean_prod = mean(production))

# Minimum and Maximum numbers for the months
#min(df_3$n) ; max(df_3$n)

# Histogram of the mean oil production
#ggplot(df_3) + geom_histogram(aes(x=mean_prod)) + xlab("Mean production (BBL/Month)")

Month_to_No = function(month.vec){
  
  # Input: H dimensional vector for characterized "month" from "Jan" to "Dec"  
  H = length(month.vec) 
  if (H > 1){
    month.output = c()
    
    for (h in 1:H){
      
      month = month.vec[h]
      
      {
        if (month == "Jan"){
          res = 1
        }
        if (month == "Feb"){
          res = 2
        }
        if (month == "Mar"){
          res = 3
        }
        if (month == "Apr"){
          res = 4
        }
        if (month == "May"){
          res = 5
        }
        if (month == "Jun"){
          res = 6
        }
        if (month == "Jul"){
          res = 7
        }
        if (month == "Aug"){
          res = 8
        }
        if (month == "Sep"){
          res = 9
        }
        if (month == "Oct"){
          res = 10
        }
        if (month == "Nov"){
          res = 11
        }
        if (month == "Dec"){
          res = 12
        }
      }
      month.output[h] = res
    }
  }
  
  if (H ==1){
    month.output = c()
    {
      if (month == "Jan"){
        res = 1
      }
      if (month == "Feb"){
        res = 2
      }
      if (month == "Mar"){
        res = 3
      }
      if (month == "Apr"){
        res = 4
      }
      if (month == "May"){
        res = 5
      }
      if (month == "Jun"){
        res = 6
      }
      if (month == "Jul"){
        res = 7
      }
      if (month == "Aug"){
        res = 8
      }
      if (month == "Sep"){
        res = 9
      }
      if (month == "Oct"){
        res = 10
      }
      if (month == "Nov"){
        res = 11
      }
      if (month == "Dec"){
        res = 12
      }
    }
    month.output = res
  }
  
  return(month.output)
}

df_4 = df_2 %>% 
  filter(production > 0) %>% 
  group_by(API10_char) %>%
  arrange(API10_char,t) %>% 
  separate(t, c("year","month"), sep = "-") %>%
  mutate(monthnumber = Month_to_No(month)) %>%
  arrange(API10_char,year,monthnumber) %>% 
  unite("year_month",year,monthnumber, sep = "-") 

extract_year_month = function(year_month_vec){
  
  # Input: H dimensional vector for characterized by the form like "17-03"
  # Output: H dimensional vector with numbers "1703"   

  H = length(year_month_vec) 
  year_month_output = c()
  
  for (h in 1:H){
    
    year_month = year_month_vec[h]
    matches <- regmatches(year_month, gregexpr("[[:digit:]]+", year_month))
    res = as.numeric(unlist(matches))[1]*100 + as.numeric(unlist(matches))[2]
    year_month_output[h] = res
  }
  
  return(year_month_output)  
}

df_5 = df_4 %>% 
  mutate(year_month_no = extract_year_month(year_month))

# Visualization
{
  Tot.N = length(unique(df_5$API10)) 
  production.ts.list = list()
  type.indicator = c() # Type 1: descling from the initial production / # Type 2: delayed declining
  
  for (i in 1:Tot.N){
    
    temp.df = df_5 %>% filter(API10 == unique(df_5$API10)[i])
    starting_year = 2000 + (temp.df$year_month_no[1] - temp.df$year_month_no[1]%%100)/100
    starting_month = temp.df$year_month_no[1]%%100
    
    production.ts.list[[i]] = ts(temp.df$production, 
                                 frequency = 12, 
                                 start = c(starting_year, starting_month))
    
    if (length(intersect(i, c(1:N))) == 1){
      # Type 1
      type.indicator[i] = 1 # Training wells
    }
    if (length(intersect(i, c(1:N))) != 1){
      # Type 2
      type.indicator[i] = 2 # Test wells
    }
  }
  
  par(mar= c(5.1, 4.5, 4.1, 2.1),
      cex.axis=1,
      cex.lab=1,
      cex.main=1,
      cex.sub=1)
  
  plot(production.ts.list[[1]], xlim = c(2011.1, 2017.6), lty = 2, ylim = c(0, max(df_5$production)), 
       ylab = c("Production (BBL/Month)"), xlab = c("Year"), col = "gray44")
  
  for (i in 2:Tot.N){
    
    if (type.indicator[i] == 1){
      points(production.ts.list[[i]], xlim = c(2011.1, 2017.6), lty = 2, type = "l", 
             col = "gray44")
    }
    if (type.indicator[i] == 2){
      points(production.ts.list[[i]], xlim = c(2011.1, 2017.6), type = "l", lty = 2, 
             col = "red")
    }
  }
  
  legend("topright", legend = c("Training wells","Testing wells"),
         col = c("gray44","red"),
         lty = c(2,2), lwd = c(2,2), cex = 1.1)
  
  title("Production rate trajectories for the training and testing wells")
}
```


Training the Spatial Weibul Model with Gibbs sampler (This can takes a time)
```{r include=FALSE}
Post.Inf.Res = SWM(Y = Y, X = X, Loc = Loc,burn=500,nmc=500,thin=10)
```


```{r}
theta.1 = Post.Inf.Res$thinned.theta.1
theta.2 = Post.Inf.Res$thinned.theta.2
theta.3 = Post.Inf.Res$thinned.theta.3
sigma.sq = Post.Inf.Res$thinned.sigma.sq
alpha.1 = Post.Inf.Res$thinned.alpha.1
alpha.2 = Post.Inf.Res$thinned.alpha.2
alpha.3 = Post.Inf.Res$thinned.alpha.3
beta.1 = Post.Inf.Res$thinned.beta.1
beta.2 = Post.Inf.Res$thinned.beta.2
beta.3 = Post.Inf.Res$thinned.beta.3
sigma.sq.1 = Post.Inf.Res$thinned.sigma.sq.1
sigma.sq.2 = Post.Inf.Res$thinned.sigma.sq.2
sigma.sq.3 = Post.Inf.Res$thinned.sigma.sq.3
gamma.sq.1 = Post.Inf.Res$thinned.gamma.sq.1
gamma.sq.2 = Post.Inf.Res$thinned.gamma.sq.2
gamma.sq.3 = Post.Inf.Res$thinned.gamma.sq.3
rho.1 = -4 ; rho.2 = -4 ; rho.3 = -4 
setwd("C:/Users/seyoonlee/OneDrive/Documents/GitHub/SWM/R codes/")
source("Prediction_SWM.R")
```

Spatial prediction for a testing well
```{r}
i.t = 346

map('state', 'texas',fill = TRUE,col = "grey",
    xlim = c(-100.5,-96),exact = TRUE,
    ylim = c(28,30))
title("Locations of training wells (black dots) and testing well (red triangle).")
points( x = Loc[,2], y = Loc[,1],  col = "black", pch = 20)
points( x = Production_Results_Eagle_Ford$Loc[i.t,2],
        y = Production_Results_Eagle_Ford$Loc[i.t,1],
        col = "red", pch = 17)
Prediction_SWM(index.of.test.well = i.t, log_scale = FALSE)
Prediction_SWM(index.of.test.well = i.t, log_scale = TRUE)
```

Spatial prediction for a testing well
```{r}
i.t = 345

map('state', 'texas',fill = TRUE,col = "grey",
    xlim = c(-100.5,-96),exact = TRUE,
    ylim = c(28,30))
title("Locations of training wells (black dots) and testing well (red triangle).")
points( x = Loc[,2], y = Loc[,1],  col = "black", pch = 20)
points( x = Production_Results_Eagle_Ford$Loc[i.t,2],
        y = Production_Results_Eagle_Ford$Loc[i.t,1],
        col = "red", pch = 17)
Prediction_SWM(index.of.test.well = i.t, log_scale = FALSE)
Prediction_SWM(index.of.test.well = i.t, log_scale = TRUE)
```
